---
title: "Uber-Data-Analysis"
author: "Vinu Baburaj"
date: "`r Sys.Date()`"
output: html_document
---


Loading required packages for this project

```{r LoadPackages, warning=FALSE, message=FALSE}
# Package names required for this project

packages <- c('ggplot2','tidyverse')

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Loading Packages
invisible(lapply(packages, library, character.only = TRUE))
```


Reading dataset csv files

```{r ReadCSV, warning=FALSE, message=FALSE}
apr <- read_csv("dataset/uber-raw-data-apr14.csv")
may <- read_csv("dataset/uber-raw-data-may14.csv")
jun <- read_csv("dataset/uber-raw-data-jun14.csv")
jul <- read_csv("dataset/uber-raw-data-jul14.csv")
aug <- read_csv("dataset/uber-raw-data-aug14.csv")
sep <- read_csv("dataset/uber-raw-data-sep14.csv")
```


Understanding fields in data

```{r CheckData}
head(apr)
head(may)
head(jun)
head(jul)
head(aug)
head(sep)
```


Merging all month's data into a single data frame

```{r JoinData}
data <- rbind(apr,may, jun, jul, aug, sep)
```


Checking the first and last few rows of merged data

```{r}
head(data)
tail(data)
```

Converting Date/Time column to a datetime object

```{r}
data$`Date/Time` <- strptime(data$`Date/Time`, "%m/%d/%Y %H:%M:%S")
```

Separating column for Day, Month, Month in words and Hour 

```{r}
data$Day <- as.numeric(format(data$`Date/Time`, format="%d"))
data$Month <- as.numeric(format(data$`Date/Time`, format="%m"))
data$MonthAbbr <- (format(data$`Date/Time`, format="%b"))
data$Hour <- as.numeric(format(data$`Date/Time`, format="%H"))
```


 Analyzing which month has the highest Uber rides booked:

```{r}
ggplot(data, aes(x=reorder(MonthAbbr,Month), fill=MonthAbbr))+
  geom_bar(color="black")+
  labs(x="Month",y="No. of Rides")+
  theme_minimal()
```

Analyzing which days have the most Uber rides booked:

```{r}
ggplot(data, aes(x=Day))+
  geom_bar()+
  labs(y="No. of Rides")+
  theme_minimal()
```

Analyzing the trends of Uber rides booked for each month:

```{r}
ggplot(data, aes(x=Day))+
  geom_bar()+
  labs(y="No. of Rides every month", x="Month")+
  theme_minimal()+
  facet_wrap(~MonthAbbr)
```
